Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-14 Manuscript/manuscript.tex
Updated "Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to the POPC membrane"

Simple binding models have been used previously to interpret
the same experimental data \cite{altenbach84,macdonald87} 
that are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
Results from measurements of the lipid headgroup order parameters and from
atomic absorption spectroscopy were best
explained by a ternary complex binding model with a binding
stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
However, a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
also provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
Note that the results based on these binding models and the measured headgroup order parameter changes
are also inluded in the tabulated experimental binding constants~\cite{marsh13}.

Since the same experimental data are reproduced by the present ECC-lipids model,
we can use the model to provide a more detailed interepration of the binding stoichiometry than
the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
complexes with various stoichiometries by counting the individual lipid
molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion,
as already done in the previous section~\ref{sec:affinity}.
The results from POPC simulation with a 285~mM concetration of CaCl$_2$ in
Fig.~\ref{fig:cacl_complexes} demonstrate the largest propensity (41\%) for the
ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
the probabilities of complexes with the stoichimetries of 1:1~(25\%)~and
1:3~(34\%)~are only slightly lower. This
suggests a more complex binding model than the previous simple ternary complex model.
%The observed propensities of
%$41\%$ for two lipids,
%$25\%$ for one lipid, and
%$34\%$ for three lipids 
With a broad brushstroke, the simulation data can, however, be viewed such that
one calcium binds to two lipids on average,
because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other 
(and complexes with more than three lipids per one calcium ion were not observed).  
This probably explains why the simple the ternary complex model fits the
experimental data, as well as the ECC-lipids simulation results, rather well
(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details).

>>>>

Simple binding models have been used previously to interpret the experimental data \cite{altenbach84,macdonald87} we use in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}). In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}. However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2}low concentrations~\cite{macdonald87}. \todo{HECTOR: Why is the following sentence remotely significant. I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.

In this work, we reproduce the same experimental data used to calculate binding stoichiometry with our ECC-POPC model. Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}. To evaluate the relative propensities for each stoichiometric complex (e.i., 1 Ca$^{2+}$: n POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.  In Fig.~\ref{fig:cacl_complexes}, from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ we found that the largest propensity is for the 1:2 complex (41\%). \todo {I seriously doubt that this is what they call a ternary complex but correct me if you know better}. Still, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower. This variability suggests a more complex binding model than the previously considered in a simple ternary complex model. With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed). This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S\ref{fig:cacl-bind} in SI).
